---
title: "Dissertation Proposal"
author: "Elizabeth Jane Wesley"
output:
  pdf_document:
    df_print: tibble
  html_notebook: default
  html_document:
    df_print: paged
---

```{r, message = F, warning = F}
library(tidyverse)
library(lubridate)
library(here)
#library(tibbletime)
library(RcppRoll)
```

## Load data.

```{r, message = F, warning = F}
read_csv(here("raw_data", "LongTermWeather.csv")) %>%
  group_by(NAME) %>% 
  summarise(min = min(DATE),
            max = max(DATE))

## Remove leap days and calculate time (day) stamp

atmo <- read_csv(here("raw_data", "LongTermWeather.csv")) %>%  
  filter(NAME == "KANSAS CITY DOWNTOWN AIRPORT, MO US") %>% 
  complete(DATE = seq.Date(min(DATE), max(DATE), by="day")) %>%
  select(TMAX, TMIN, DATE) %>% 
  filter(!(month(DATE) == 2 & day(DATE) == 29)) %>% 
  gather(key = "Var", value = "Value", -DATE) %>% 
  group_by(year(DATE), Var) %>% 
  mutate(time = row_number()) %>% 
  ungroup() %>% 
  select(-"year(DATE)")
```

# There are NA values for 4 dates within the asthma date range. Let's look at the surrounding dates to see if it would be appropriate to inpute a mean

```{r}
# miss.2005 <- asthma %>% 
#   filter(date > "2005-01-07", date < "2005-01-14")
# miss.2012 <- asthma %>% 
#   filter(date > "2012-02-25", date < "2012-03-03")
# miss.2011 <- asthma %>% 
#   filter(date > "2011-09-12", date < "2011-09-18")
```

The answer is no!! 

# Inpute values for NA dates from the Leavenworth station
# https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USW00003947/detail

```{r}
inpute <- tribble(
  ~ TMAX, ~ TMIN, ~ DATE,
  32, 25, "2005-01-10",
  30, 24, "2005-01-11",
  59, 40, "2011-09-15",
  61, 39, "2012-02-28") 

inpute <- inpute %>% 
  mutate(DATE = as_date(DATE),
         TMAX = round(((TMAX - 32) * 5/9), 1), 
         TMIN = round(((TMIN - 32) * 5/9), 1)) %>% 
  gather("Var", "Value", -DATE)

atmo <- atmo %>% 
  left_join(inpute, by = c("DATE", "Var")) %>% 
  mutate(Value = coalesce(Value.x, Value.y)) %>% 
  select(-Value.x, -Value.y)
```

## Percentiles across all years within a five-day window

```{r}
atmo2 <- tibble()

# cumulative distance within a five day rolling window; group_by(Var) did something weird with the cume_dist
# window function so I used a for loop instead of map
# There are some missing values which are ignored

for (v in c("TMAX", "TMIN")){
  
  var <- filter(atmo, Var == v) 
  cdfs <- tibble()
  
  for (d in unique(atmo$time)){
    
    cdf <- bind_rows(filter(var, time == d - 2),
                    filter(var, time == d - 1),
                    filter(var, time == d),
                    filter(var, time == d + 1),
                    filter(var, time == d + 2)) %>% 
           arrange(Value) %>% 
           mutate(cdf = cume_dist(Value)) %>% 
           filter(time == d)
    cdfs <- bind_rows(cdfs, cdf)
    
  }
  atmo2 <- bind_rows(atmo2, cdfs)
}
```


```{r}
## Test the cdf calculation

# test <- atmo2 %>% filter(Var == "TMIN", time == d - 2)
# test2 <- atmo2 %>% filter(Var == "TMIN", time == d -1)
# test3 <- atmo2 %>% filter(Var == "TMIN", time == d)
# test4 <- atmo2 %>% filter(Var == "TMIN", time == d + 1)
# test5 <- atmo2 %>% filter(Var == "TMIN", time == d + 2)
# t <- bind_rows(test, test2, test3, test4, test5)
# t <- t %>%  arrange(Value)
# t <-  t %>% mutate(cdf.test = cume_dist(Value)) %>% 
#   filter(time == d)
# 
# sum(t$cdf != t$cdf.test, na.rm = T)
# 
# ggplot(filter(atmo2, time == 180)) + 
#   geom_point(aes(Value, cdf)) + 
#   facet_wrap(~ Var, ncol = 1, scales = "free")
```

## Join asthma data and make long, right join to keep both TMAX and TMIN, drop all rows where there are not counts associated with the temperature variables

```{r}
asthma <- read_csv(here("raw_data", "asthma_noid.csv")) %>% 
  select(date, count) %>% 
  distinct() %>% 
  arrange(count) %>% 
  mutate(count.cdf = cume_dist(count)) %>% 
  right_join(atmo2, by = c("date" = "DATE")) %>% 
  mutate(year = year(date),
          month = month(date),
          day = wday(date, label = T, week_start = getOption("lubridate.week.start", 1)),
          doy = yday(date)) %>% 
  drop_na(count)
```

# Asthma counts show a clear increasing trend across years. I think this is probably not a real increase but due to reporting or something like that. Remove the linear trend.

```{r}
asthma.fit <- lm(count ~ year, data = asthma)
summary(asthma.fit)

asthma <- asthma %>% 
  mutate(count.detrend.yr = asthma.fit$residuals,
         count.detrend.yr.cdf = cume_dist(count.detrend.yr))
```

# Number of days in a row the atmo variables are above the 80th, 90th, 95th, & 99th percentiles, or below the 1st, 5th, 10th, or 20th

```{r}
asthma2 <- asthma

h.pcts <- c(.80, .90, .95, .99)
l.pcts <- c(.01, .05, .10, .20)

for (p in h.pcts){
  
  tmax <- asthma %>% 
    filter(Var == "TMAX") %>% 
    mutate(pct = 0) %>% 
    arrange(date)
    
  for (row in 1:nrow(tmax)){
    if (row == 1 & tmax$cdf[row] >= p){
      tmax$pct[row] = 1
    } else if (tmax$cdf[row] < p){
      tmax$pct[row] = 0
    } else {
      tmax$pct[row] = tmax$pct[row - 1] + 1
    }
  }
  
  tmin <- asthma %>% 
    filter(Var == "TMIN") %>% 
    mutate(pct = 0) %>% 
    arrange(date)
  
  for (row in 1:nrow(tmin)){
    if (row == 1 & tmin$cdf[row] >= p){
      tmin$pct[row] = 1
    } else if (tmin$cdf[row] < p){
      tmin$pct[row] = 0
    } else {
      tmin$pct[row] = tmin$pct[row - 1] + 1
    }
  }
  
  x <- bind_rows(tmax, tmin) %>% 
    rename(!!sym(paste0("pct", p)) := pct)
  
  
  
  asthma2 <- left_join(asthma2, x)
}

for (p in l.pcts){
  
  tmax <- asthma %>% 
    filter(Var == "TMAX") %>% 
    mutate(pct = 0) %>% 
    arrange(date)
    
  for (row in 1:nrow(tmax)){
    if (row == 1 & tmax$cdf[row] <= p){
      tmax$pct[row] = 1
    } else if (tmax$cdf[row] > p){
      tmax$pct[row] = 0
    } else {
      tmax$pct[row] = tmax$pct[row - 1] + 1
    }
  }
  
  tmin <- asthma %>% 
    filter(Var == "TMIN") %>% 
    mutate(pct = 0) %>% 
    arrange(date)
  
  for (row in 1:nrow(tmin)){
    if (row == 1 & tmin$cdf[row] <= p){
      tmin$pct[row] = 1
    } else if (tmin$cdf[row] > p){
      tmin$pct[row] = 0
    } else {
      tmin$pct[row] = tmin$pct[row - 1] + 1
    }
  }
  
  x <- bind_rows(tmax, tmin) %>% 
    rename(!!sym(paste0("pct", p)) := pct)
  
  
  
  asthma2 <- left_join(asthma2, x)
}
```
# Make long

```{r}
asthma.long <- asthma2 %>% 
  gather(key = "Percentile", value = "Run", starts_with("pct"))
```

## Plot variable by date.

```{r}
# Counts show a consistent pattern as a function of weekday. Mondays are highest declining to a low on Saturday.
ggplot(asthma2) +
  geom_col(aes(day, count.detrend.yr, fill = day)) +
  facet_wrap(~ year, scales = "free")

asthma.long %>% filter(Var == "TMAX", Run != 0) %>% 
  ggplot() +
  geom_point(aes(Run, count.detrend.yr)) +
  geom_smooth(aes(Run, count.detrend.yr), method = "lm") +
  facet_wrap(~ Percentile, scales = "free")

asthma.long %>% 
  filter(Var == "TMAX") %>% 
  ggplot() +
  geom_point(aes(Value, count.detrend.yr)) +
  geom_smooth(aes(Value, count.detrend.yr), method = "lm") +
  facet_wrap(~ month, scales = "free")

asthma2 %>% 
  filter(Var == "TMAX") %>% 
  ggplot() +
  geom_point(aes(Value, count.cdf)) +
  geom_smooth(aes(Value, count.cdf), method = "lm") +
  facet_wrap(~ month, scales = "free_x")

asthma2 %>% 
  filter(Var == "TMAX") %>% 
  ggplot() +
  geom_point(aes(Value, count.detrend.yr.cdf)) +
  geom_smooth(aes(Value, count.detrend.yr.cdf), method = "lm") +
  facet_wrap(~ month, scales = "free_x")

asthma2 %>% 
  filter(Var == "TMAX") %>% 
  ggplot() +
  geom_point(aes(cdf, count.detrend.yr.cdf)) +
  geom_smooth(aes(cdf, count.detrend.yr.cdf), method = "lm") +
  facet_wrap(~ month, scales = "free_x")

asthma2 %>% 
  filter(Var == "TMAX", pct0.8 > 0) %>% 
  ggplot() +
  geom_point(aes(cdf, count.detrend.yr.cdf)) +
  geom_smooth(aes(cdf, count.detrend.yr.cdf), method = "lm") +
  facet_wrap(~ month, scales = "free_x")

asthma2 %>% 
  filter(Var == "TMAX", count.detrend.yr.cdf >= 0.8) %>% 
  ggplot() +
  geom_point(aes(Value, count.detrend.yr.cdf)) +
  geom_smooth(aes(Value, count.detrend.yr.cdf), method = "lm") +
  facet_wrap(~ month, scales = "free")
  
```

# t-test for TMAX percentile counts vs TMAX non percentile counts

```{r}
diff.dist <- function(pct){
  
  high <- asthma.long %>% 
    filter(Var == "TMAX", month %in% c(7, 8), cdf >= pct) %>% 
    select(count) %>% 
    unlist()
  
  not <-  asthma.long %>% 
    filter(Var == "TMAX", month %in% c(7, 8), cdf < pct) %>% 
    select(count) %>% 
    unlist()
  
  t.test(high, not)
  
  ggplot() +
    geom_density(aes(high, color = "red")) +
    geom_density(aes(not, color = "blue")) +
    scale_color_manual(labels = c("Exceeds TMAX\npercentile", "Below TMAX\npercentile"), values = c("red", "blue")) +
    labs(title = pct) +
    xlab("Count")
}

map(h.pcts, diff.dist)

```

# t-test for count percentiles  vs count non percentile 

```{r}
diff.dist <- function(pct){
  
  high <- asthma.long %>% 
    filter(Var == "TMAX", month %in% c(7, 8), count.detrend.yr.cdf >= pct) %>% 
    select(Value) %>% 
    unlist()
  
  not <-  asthma.long %>% 
    filter(Var == "TMAX", month %in% c(7, 8), count.detrend.yr.cdf < pct) %>% 
    select(Value) %>% 
    unlist()
  
  print(t.test(high, not))
  
  ggplot() +
    geom_density(aes(high, color = "red")) +
    geom_density(aes(not, color = "blue")) +
    scale_color_manual(labels = c("Exceeds count\npercentile", "Below count\npercentile"), values = c("red", "blue")) +
    labs(title = pct) +
    xlab("TMAX")
}

map(h.pcts, diff.dist)

```
# 99th percentile

```{r}
asthma.long %>% 
  filter(Var == "TMAX", cdf >= 0.99) %>% 
  ggplot() +
  geom_point(data = asthma.long, aes(date, count)) +
  geom_point(aes(date, count), color = "red", size = 2) 
  


```


