# Data {#data}

## Study area {#study-area}

The Kansas City metropolitan area as delineated by the United States Census Bureau is located at 39.0398$^{\circ}$N latitude and 94.5949$^{\circ}$W longitude and spans two states and six counties: Johnson and Wyandotte Counties in Kansas, and Platte, Clay, Cass, and Jackson Counties in Missouri. The Köppen climate classification is humid subtropical (Cfa), with rainfall year round, averaging 964 mm annually. The annual temperature average is 12.8$^{\circ}$C, with a maximum average high of 26.1$^{\circ}$C in July and a minimum average low of -2$^{\circ}$C in January (https://en.climate-data.org/location/715044). The Kansas City metro area exhibits characteristic patterns of urban sprawl (Fig. \@ref(fig:KC)), which is generally defined as “geographic expansion over large areas, low-density land use, low land-use mix, low connectivity, and heavy reliance on automobiles relative to other modes of travel”  [@Stone2010] showing a 55\% increase in built area between 1972 and 2001 [@Ji2008]. The Kansas City metro area had an estimated population of 2,142,419 in 2018, a 5\% increase from 2000. 

\begin{table}
	\caption{Demographic characteristics of the pediatric population of the Kansas City metro area. 2018 American Community Survey (https://data.census.gov/cedsci) \label{tab:demo-data}}
	\centering
	\begin{tabular}{lll}
		\toprule
		Category & Attribute & Percent \\
		\midrule
		Age   & Pop. under 18 & 24.2\% \\
		\midrule
		Race  & White alone & 73.9\% \\
		      & Black alone & 7.4\% \\
		      & Asian alone & 3.2\% \\
		      & Native Hawaiian alone & 0.4\% \\
		      & American Indian alone & 0.3\% \\
		      & Some other race alone & 2.6\% \\
		      & Two or more races & 6.9\% \\
		      & Hispanic or latino of any race & 11.7\% \\
		\midrule
		Other & Household below poverty level & 5.1\% \\
		      & Household receiving public assistance & 7.6\% \\
		      & Single-parent household & 31.4\% \\
		\bottomrule
	\end{tabular}
\end{table}

```{r KC, fig.cap=cap, fig.pos="h"}
include_graphics("images/KC.pdf")

cap <- "The Kansas City metro area exhibits classic patterns of urban sprawl. True-color composite image from Landsat (30 m) June 6, 2011."
```


## Pediatric asthma {#asthma-data}

KC Health CORE is a collaborative initiative between Children’s Mercy Hospital (CMH) and the Center for Economic Information (CEI) at the University of Missouri, Kansas City created to investigate the geographic disparity of pediatric health outcomes. This analysis will use pediatric asthma data from 2001-2012 geocoded to street centerlines based on the patients’ home address at the time of admission (Fig. \@ref(fig:asthma)). These observations are readily aggregated to counts or rates per space-time interval. Although the dataset contains information on the years 1999 and 2001, these observations are inconsistent with the remaining years, likely due to differences in data procedures and will not be considered. More data will be made available during the course of this research with observations through 2019 that will be incorporated as is possible. The data comes from a retrospective collection of pediatric asthma encounters within the CMH network. Only children ages 2-18 are considered as asthma is difficult to diagnose in very young children [@kaneInterdisciplinaryHealthDisparities2020; @trivediAsthmaChildrenAdults2019]. The original medical records were formatted according to Table \@ref(tab:original-asthma-data) [@kaneInterdisciplinaryHealthDisparities2020].

\begin{table}
	\caption{Structure of the original pediatric asthma data records submitted by CMH to UMKC-CEI. \label{tab:original-asthma-data}}
	\centering
	\begin{tabular}{ll}
		\toprule
		Category & Attributes \\
		\midrule
		Diagnosis & Date of admission \\
		& ICD-9 code \\
		& Event account number \\
		& Patient medical record number (MRN) \\
		& Patient residential address \\
		\midrule
		Demographics & Birthdate \\
		& Sex \\
		& Race \\
		& Ethnicity \\
		\midrule
		Visit characteristics & Payment type \\
		& Patient class \\
		\bottomrule
	\end{tabular}
\end{table}

The data were further classified into three severity levels according to the International Classification of Diseases, 9^th^ revision diagnoses codes (ICD-9) and the patient class: (1) controlled visit, (2) acute care visit, and (3) hospitalization. The patient class records both the location and the type of treatment received by the patient–e.g. controlled vs. acute care, inpatient vs. outpatient, etc. Level-1 visits are controlled encounters that do not necessarily indicate presentation of symptoms but could be used to identify the general pediatric asthma population. Level-2 visits are acute care visits for immediate symptoms at same-day clinics and emergency departments. Level-3 visits are the most severe, requiring hospitalization and/or aggressive treatment. The CMH network---while serving all children within the Kansas City Metropolitan Area---is not likely to capture all of the pediatric population. However, as the only facilities that specialize in pediatric care, the network is likely to capture the majority of the acute and most severe cases as emergency patients are typically transferred there from other facilities. See @kaneInterdisciplinaryHealthDisparities2020 for more details and to read her excellent work.

```{r asthma, fig.cap=cap, fig.pos="h"}
include_graphics("images/asthma_heatmap.pdf")

cap <- "Asthma observations from 2002 to 2012. Points are partially transparent to both show spatial density of points and to preserve patient privacy."
```


## Climatic data {#atmo-data}

Climatic data include both satellite and station observations. The spatial variation of climatic data will be assessed using remotely-sensed data including LST (Fig. \@ref(fig:LST)) and other variables to be determined based on availability. LST from the Moderate Resolution Imaging Spectroradiometer (MODIS) are available daily at 1 km spatial resolution, with two passovers during the daytime---once in the morning and once in the afternoon---and two during the nighttime. Cloud cover produces low-quality images so the 8-day composite images may be a suitable alternative depending on the desired temporal resolution. It is important to note however, that the composite images bias LST and increase the magnitude of the surface urban heat island (SUHI), impacts which vary seasonally [@Hu2013]. Remotely-sensed imagery can be accessed and manipulated using the Google Earth Engine analysis platform (https://earthengine.google.com/). Derived variables may include LST anomalies relative to the mean and the diurnal temperature range. 

Station data were retrieved from the NOAA National Centers for Environmental Information for the Kansas City Downtown Airport, MO, US. The station is located at 39.1208$^{\circ}$N, 94.5969$^{\circ}$W. Daily precipitation totals, maximum temperature, and minimum temperature were retrieved for all dates between 01-01-1900 and 10-19-2019. Daily average wind speed, direction of fastest 2-minute wind, and direction of fastest 5-minute wind were retrieved for the years 2000-2012. Daily maximum 8-hour O$_3$ concentration and daily mean PM$_{2.5}$ concentration were retrieved from the EPA for the JFK Community Center in Kansas City, KS, US, located at 39.117219$^{\circ}$N, 94.635605$^{\circ}$W for the years 2000-2012. These data represent the daily temporal variability of climatic conditions for the Kansas City metro area but do not describe any spatial variability. Derived variables may include daily percentile values, number of days in a row that values exceed a threshold (e.g. 95$^{th}$ percentile maximum temperature days to indicate a heat wave), diurnal ranges, and seasonal trends.

```{r LST, fig.cap=cap, fig.pos="h"}
include_graphics("images/LST.pdf")

cap <- "LST calculated from Landsat (30 m) June 6, 2011 shows the UHI in Kansas City."
```


## Environmental data {#environ-data}

High-resolution land-cover data comes from the Mid-America Regional Council (MARC). MARC created the Natural Resources Inventory (NRI) map of Greater Kansas City with an object-based classification, using SPOT data from May, June, and August of 2012 as well as ancillary data (LiDAR, hydrography, parcels/zoning class, transportation centerlines, streamlines, and floodplains). The resulting land cover map has an estimated accuracy of 83 - 91\% for the Level I classifications of impervious, barren, vegetated, and water. Impervious comprises buildings and other impervious surfaces, barren comprises land with 0 - 10\% vegetated fraction, vegetated comprises land with 10-100\% vegetated fraction, and water comprises water features. The spatial resolution of the NRI land cover map is 2.5 m and the extent is the 4,423 square miles that comprise the 9 county Kansas City metropolitan area [@Mid-2013]. An alternative or complementary land cover dataset is the National Land Cover Dataset (NLCD) produced by the United States Geological Survey (USGS), a 30 m resolution land-cover map based on Landsat imagery that includes the percent of impervious surface cover (http://www.mrlc.gov). A more specifically urban-oriented classification system is the local climate zone (LCZ) typology---"regions of uniform surface cover, structure, material, and human activity that span hundreds of meters to several kilometers in horizontal scale" [@stewartLocalClimateZones2012]. The scheme delineates 17 separate LCZs, each with a distinct temperature regime originating from a relatively homogeneous collection of surface properties, primarily the density and height of roughness objects [@chingWUDAPTUrbanWeather2018].

Additionally, spatial features from the built environment will be acquired for the study area, including, but not limited to, transportation networks, hazardous facilities, and power infrastructure. Traffic density data will also be acquired if possible. Derived variables may include density of or proximity to infrastructure networks or greenspace. 

```{r green, fig.cap=cap, fig.pos="h"}
include_graphics("images/vegmap.pdf")

cap <- "Binary map of vegetated and non-vegetated surfaces from the MARC NRI (2.5 m)."
```


## Social determinants of health data {#soc-data}

There are two primary levels of social determinants of health for this study: individual and community. Each observation of pediatric asthma is coded for sex, race, age, ethnicity, and payment type---specifically Medicaid or other state coverage, private insurance, or self pay. Community-level characteristics come from the United States Census (https://data.census.gov/cedsci/) (Fig. \@ref(fig:med-income)). Variables include the percent of minority residents, percent of households below the poverty level, the percent of owner-occupied households, and the number of children between 2 and 18 years of age. While the block group is the highest resolution of census geography for which this information is available, the unit of analysis will likely be the census tract due to the geographic sparsity of daily pediatric asthma observations. The census tract is roughly equivalent to a neighborhood and the associated data provide an appropriate proxy measure for community structure.

```{r med-income, fig.cap=cap, fig.pos="h"}
include_graphics("images/med_income.pdf")

cap <- "Median income per census tract from the 2010 United States Decennial Census."
```

